
---
title: "Medicaid Outpatient Narcotics"
output: html_notebook
---

This is an [Dataset](https://data.medicaid.gov/State-Drug-Utilization/State-Drug-Utilization-Data-2016/3v6v-qk5s) Notebook. When you execute code within the notebook, the results appear beneath the code. 

This is a [description of the dataset](file:///C:/Users/Chris%20Iyer/Downloads/Drug_Utilization_Data_Descriptions.pdf)

```{r}
library(knitr)
library(dplyr)
library(rvest)
library(ggplot2)
library(gridExtra)
```


###Read in Data

```{r}
medicaid <- read.csv("medicaid.csv", header = T, stringsAsFactors = F)
```

###Retrieve narcotics list

```{r}
Narcotics <- read_html("https://www.drugs.com/drug-class/narcotic-analgesics.html")
Drug_Names <- Narcotics %>% html_nodes("br+ .small a") %>% html_text()

```

Population Data

```{r}
Pops <- read_html("http://worldpopulationreview.com/states/")
StateName <- Pops %>% html_nodes("td:nth-child(2)") %>% html_text()
#StateName
StatePopul <- Pops %>% html_nodes("td~ td+ td") %>% html_text()
#StatePopul
StatePopul <- as.numeric(gsub(",", "", StatePopul))
Population <- cbind.data.frame(StateName, StatePopul)
#str(Population)
Population
Population["State"] <- state.abb[match(Population$StateName,state.name)]
Population

```


###Convert medicaid data to a dataframe

```{r}
Medicaid <- as.data.frame(medicaid)
str(Medicaid)

kable(head(Medicaid))

```
###Filter some of the narcotics from the medicaid dataframe

```{r}

Drug_Names <- unique(toupper(Drug_Names))
Drugs <- c("MORPHINE",
"OXYMORPHONE", 
"HYDROMORPHONE", 
"BUPRENORPHINE", 
"TRAMADOL", 
"HYDROCODONE", 
"BUTORPHANOL",  
"PROPOXYPHENE", 
"LEVORPHANOL",
"TAPENTADOL",
"PENTAZOCINE",  
"REMIFENTANIL")

medicaid1 <- Medicaid %>% filter(Product.Name %in% Drug_Names)
dim(medicaid1)
unique(medicaid1$Product.Name)
MORPHINE <- Medicaid %>% filter(grepl("MORPHINE", Product.Name))
OXYMORPHONE <- Medicaid %>% filter(grepl("OXYMORPHONE", Product.Name))
HYDROMORPHONE <- Medicaid %>% filter(grepl("HYDROMORPHONE", Product.Name))
BUPRENORPHINE <- Medicaid %>% filter(grepl("BUPRENORPHINE", Product.Name))
TRAMADOL <- Medicaid %>% filter(grepl("TRAMADOL", Product.Name))
HYDROCODONE <- Medicaid %>% filter(grepl("HYDROCODONE", Product.Name))
BUTORPHANOL  <- Medicaid %>% filter(grepl("BUTORPHANOL", Product.Name))
PROPOXYPHENE <- Medicaid %>% filter(grepl("PROPOXYPHENE", Product.Name))
LEVORPHANOL  <- Medicaid %>% filter(grepl("LEVORPHANOL", Product.Name)) 
TAPENTADOL <- Medicaid %>% filter(grepl("TAPENTADOL", Product.Name))
PENTAZOCINE <- Medicaid %>% filter(grepl("PENTAZOCINE", Product.Name))
REMIFENTANIL <- Medicaid %>% filter(grepl("REMIFENTANIL", Product.Name))
```

###Bind narcotics into a dataframe

```{r}
allDrugs <- rbind.data.frame(medicaid1, MORPHINE, OXYMORPHONE, HYDROMORPHONE, BUPRENORPHINE, TRAMADOL, HYDROCODONE, BUTORPHANOL, PROPOXYPHENE, LEVORPHANOL, TAPENTADOL, PENTAZOCINE, REMIFENTANIL )
dim(allDrugs)
```

All the Narcotics in All States

```{r}
dim(allDrugs)
head(allDrugs)
allDrugs <- allDrugs %>% arrange(desc(Number.of.Prescriptions))
head(allDrugs, 35)
```
Total number of prescriptions by state, NAs eliminated

```{r}
allDrugs %>% 
  group_by(State) %>% 
  summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE), sumAmount = sum(Medicaid.Amount.Reimbursed,na.rm=TRUE))
```


Percent of NA Values per State

```{r}
allDrugsXX <- allDrugs %>% group_by(State) %>% summarise(mean(is.na(Number.of.Prescriptions)))
allDrugsXX
```


All Drugs with State XX removed, NAs eliminated

```{r}
States <- allDrugs %>% filter(State != "XX") %>% 
  group_by(State) %>% 
  summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE), sumAmount = sum(Medicaid.Amount.Reimbursed,na.rm=TRUE))
States
head(allDrugs)
allDrugs <- inner_join(allDrugs, Population)
allDrugs
allDrugsnoXX <- inner_join(States, Population)
allDrugsnoXX
```

Number of prescriptions per capita

```{r}
PerPerson <- allDrugsnoXX %>% mutate(RxPP = sumNumber/StatePopul, RxSpendingPP = sumAmount/StatePopul) %>% arrange(desc(RxPP))
```


```{r}
OpiodDeath <- read.csv("OpiodDeath.csv", header = T)
OpiodDeath <- OpiodDeath[c(1:51), c(1,4)]
OpiodDeath$Number <- as.numeric(gsub(",", "", OpiodDeath$Number))
OpiodDeath 
```

Opiod Death per capita

```{r}
DeathPerCap <- inner_join(PerPerson, OpiodDeath)
DeathPerCap <-  DeathPerCap %>% mutate(ODPP = Number/StatePopul)
DeathPerCap

```


```{r}
ODperCap <- ggplot(DeathPerCap, aes(x = State, y = ODPerCap, color = State, fill = State)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
RxPerP <- ggplot(DeathPerCap, aes(x = State, y = RxPP, color = State, fill = State)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
grid.arrange(ODperCap, RxPerP, ncol=2)

```

where you subtract the mean and divide by the standard deviation of your variable

```{r}

meanOD <- mean(DeathPerCap$ODPerCap)
meanOD
sdOD <- sd(DeathPerCap$ODPerCap)
sdOD

```

```{r}
meanRx <- mean(DeathPerCap$RxPP)
meanRx
sdRx <- sd(DeathPerCap$RxPP)
sdRx
```

where you subtract the mean and divide by the standard deviation of your variable

```{r}
names(DeathPerCap)
DeathandRxperCapperState <- DeathPerCap 
names(DeathandRxperCapperState)
DeathandRxperCapperState <- DeathandRxperCapperState %>% mutate(zRx = ((RxPP - mean(RxPP))/ sd(RxPP)), zOD = ((ODPP - mean(ODPP))/sd(ODPP)))
DeathandRxperCapperState
```
```{r}
names(DeathandRxperCapperState)


library(reshape2)


newset <- DeathandRxperCapperState %>% select(State, zRx, zOD)
newset
dat.m <- melt(newset)
range(dat.m$value)
dat.m

newset

ggplot(dat.m, aes(State, value, fill=variable)) + 
  geom_bar(position="dodge",stat="identity")


```


forCorr <- DeathPerCap %>% select(State, RxPP, ODPerCap)
library(Hmisc)
forCorr
m1<-lm(log(ODPerCap) ~ RxPP,data=forCorr)
summary(m1)
plot(m1)



ggplot(DeathPerCap, aes(x=RxPP, y=ODPerCap)) +
    geom_point(shape=1)      # Use hollow circles

ggplot(DeathPerCap, aes(x=RxPP, y=ODPerCap)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm)   # Add linear regression line 
                             #  (by default includes 95% confidence region)

ggplot(DeathPerCap, aes(x=RxPP, y=ODPerCap)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm,   # Add linear regression line
                se=FALSE)    # Don't add shaded confidence region


ggplot(DeathPerCap, aes(x=RxPP, y=ODPerCap)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth()            # Add a loess smoothed fit curve with confidence region
#> `geom_smooth()` using method = 'loess'


ggplot(DeathPerCap, aes(x=RxPP, y=ODPerCap)) +  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) + geom_point()



linear.model <-lm(ODPerCap ~ RxPP, DeathPerCap)
log.model <-lm(log(ODPerCap) ~ RxPP, DeathPerCap)
exp.model <-lm(ODPerCap ~ exp(RxPP), DeathPerCap)

log.model.df <- data.frame(x = DeathPerCap$RxPP,
                           y = exp(fitted(log.model)))
log.model.df
ggplot(DeathPerCap, aes(x=RxPP, y=ODPerCap)) + 
  geom_point() +
  geom_smooth(method="lm", aes(color="Exp Model"), formula= (ODPerCap ~ exp(RxPP)), se=FALSE, linetype = 1) +
  geom_line(data = log.model.df, aes(x, y, color = "Log Model"), size = 1, linetype = 2) + 
  guides(color = guide_legend("Model Type"))




XX: What is XX?

```{r}
XX <- allDrugs %>% filter(State == "XX") %>% group_by(State) %>% 
  summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE), sumAmount = sum(Medicaid.Amount.Reimbursed,na.rm=TRUE)) 
XX
```




Is XX the total of all the states prescriptions?

```{r}
sum(States$sumNumber)/XX$sumNumber
sum(States$sumAmount)/XX$sumAmount
XX$sumNumber/sum(States$sumNumber)
XX$sumAmount/sum(States$sumAmount)
```

Quarterly number of prescriptions

```{r}
head(allDrugs)
Quarterly <- allDrugs %>% group_by(State, Quarter) %>% summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE))
```

###Quarterly Count of NA Values per state
We don't know the number of prescriptions for each NA value

```{r}
Subs1<-subset(allDrugs, (is.na(allDrugs[,11]))) 
#Total number of NA
dim(Subs1)
```

###Quartrely NA for states

```{r}
Subs2 <- Subs1%>% group_by(State, Quarter) %>% summarise(NA_count = n())
dim(Subs2)
head(Subs2, 28)
```

#Number of NA for each state in 2016

```{r}
Subs3 <- Subs1 %>% group_by(State) %>% summarise(NA_count = n())
dim(Subs3)
kable(Subs3)
```


Percent of NA entries in the allDrugs dataset. 54%

```{r}
NoNA <- allDrugs %>% na.omit() 
head(NoNA)

dim(allDrugs)
dim(NoNA)
1-(28317/62034)
```


###Arkansas

```{r}
AK <- allDrugs %>% filter(State == "AK") 
dim(AK)
#There are 826 entries in AK
sum(is.na(AK$Number.of.Prescriptions))
summary(AK)
  
head(AK)
#Percentage of values that are NA
480/826
hist(AK$Number.of.Prescriptions)
```

###Percent NA in Maine

```{r}
ME <- allDrugs %>% filter(State == "ME") 
dim(ME)
sum(is.na(ME$Number.of.Prescriptions))
summary(ME)
head(ME)
#Percent NA in Maine
293/688
hist(ME$Number.of.Prescriptions)
```

###CA

```{r}
CA <- allDrugs %>% filter(State == "CA") 
dim(CA)
sum(is.na(CA$Number.of.Prescriptions))
summary(CA)
head(CA)
#Percent of NA
1275/2904
hist(CA$Number.of.Prescriptions)
```


Visualization of missing values

```{r, fig.height=8, fig.width=10}
library(Amelia)

missmap(allDrugs, main="Missing Data Map")
```


```{r}
library(reshape2)
library(ggplot2)
library(plyr)
allDrugsNoNA <- allDrugs %>% na.omit()
head(allDrugsNoNA)
dim(allDrugsNoNA)
summary(allDrugs)
allDrugsNoNA$z <- ave(allDrugsNoNA$Number.of.Prescriptions, allDrugsNoNA$State, FUN=scale)
head(allDrugsNoNA)

new_DF <- allDrugs[rowSums(is.na(allDrugs)) > 0,]
new_DF[new_DF=='NA'] <- NA
head(new_DF)


```

```{r, eval=FALSE, fig.height=75, fig.width=10, include=FALSE}

ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions, fill = factor(State))) + geom_histogram(binwidth = 150) + stat_bin(binwidth = 100, geom="text", aes(label=..count..), angle = 90) + facet_wrap(~ State, ncol = 1) +  theme_bw() + scale_x_continuous(limits = c(0, 5000)) 

```

[labels on ggplot](https://stackoverflow.com/questions/30057765/histogram-ggplot-show-count-label-for-each-bin-for-each-category)

```{r, fig.height=45, fig.width=10}
#head(allDrugsNoNA)
ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions, group= factor(State), fill = "white")) + geom_histogram() + stat_bin(binwidth = 100, geom="text", aes(label=..count..), angle = 90) + facet_wrap(~ State, ncol = 2) + theme_bw() + scale_x_continuous(limits = c(-100, 1100), breaks = c(-100, 0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100)) +
  theme(axis.text.x=element_text(angle=90,hjust=1)) 
```


```{r, fig.height=45, fig.width=10}
p7 <- ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions)) +
        geom_histogram() + facet_wrap(~State, ncol = 2) + scale_x_continuous(limits = c(-100, 10000), breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000, 11000)) + scale_y_continuous(limits = c(-100,275)) + stat_bin(binwidth = 400, geom="text", aes(label=..count..), angle = 90)
p7
```


```{r}
library(dplyr)
head(allDrugs)
x <- allDrugs %>% group_by(state = State) %>% summarise(variable = sum(is.na(Number.of.Prescriptions))/prod(dim(Number.of.Prescriptions)))
x

y <- allDrugs %>% group_by(State) %>% dplyr::summarise(count = n())
y

z <- cbind(x,y)

z <- z %>% mutate(PercentNA = variable/count) %>% select(State, PercentNA)
z

```


```{r, fig.height=56, fig.width=10}
ggplot(allDrugsNoNA,aes(x=Number.of.Prescriptions)) + 
    stat_bin(binwidth=500, geom="text", aes(label=..count..), vjust=-1.5, angle = 90) + ylim(c(-100, 550)) +  facet_wrap(~State, ncol = 1) + scale_x_continuous(limits = c(-2000, 6000), breaks = c(-2000,-1000, 0, 500,1000, 1500,2000, 2500,3000, 3500,4000, 4500,5000) ) +
  theme(axis.text.x=element_text(angle=90,hjust=1)) 
```


```{r}
ggplot(z, aes(x = State, y = PercentNA)) + geom_bar(stat = "identity")
```


```{r, fig.height=50, fig.width=10}


ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions )) + 
  geom_histogram(aes(fill = ..count..), breaks = seq(0,1000, by = 100),binwidth=100, colour="blue", fill = "blue", alpha = 0.2, lwd=0.2) +
  stat_bin(aes(label = ..count..),geom="text", colour="black", vjust=-1, size=4.5, breaks = c(0, 100, 200,300,400,500, 600,700,800,900,1000)) +
  scale_x_continuous(limits = c(0, 1000),labels = c(0, 100, 200,300,400,500, 600,700,800,900,1000), breaks = c(0, 100, 200,300,400,500, 600,700,800,900,1000)) + scale_y_continuous(limits = c(0,500), labels = c(0, 100, 200, 300, 400, 500))+ facet_wrap(~State, ncol = 2) + theme(axis.text.x=element_text(angle=90,hjust=0)) 
```



```{r}
dat <-structure(list(sex = structure(c(1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L), .Label = c("male", 
"female"), class = "factor"), A = c(1, 2, 0, 2, 1, 2, 2, 0, 2, 
0, 1, 2, 2, 0, 0, 2, 0, 0, 0, 2), B = c(0, 0, 0, 0, 0, 2, 0, 
0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0), C = c(1, 2, 1, 0, 0, 
2, 1, 1, 0, 1, 1, 0, 1, 2, 1, 2, 0, 2, 1, 2), D = c(2, 2, 0, 
2, 2, 2, 1, 0, 1, 1, 1, 0, 1, 2, 0, 0, 1, 1, 1, 0), E = c(0, 
0, 0, 1, 0, 0, 1, 0, 0, 1, 0, 0, 0, 2, 0, 0, 0, 1, 2, 2), F = c(2, 
2, 1, 2, 1, 2, 2, 0, 1, 2, 0, 1, 2, 2, 0, 1, 2, 2, 2, 2)), .Names = c("sex", 
"A", "B", "C", "D", "E", "F"), variable.labels = structure(c("sex", 
"zenuwac", "panieke", "gespann", "rustelo", "angstig", "onzeker"
), .Names = c("sex", "anx01", "anx02", "anx03", "anx04", "anx05", 
"anx06")), codepage = 20127L, row.names = c(NA, 20L), class = "data.frame")
str(dat)
dat
do.call(grid.)
```

[How to do a frequency table in base R](http://www.r-tutor.com/elementary-statistics/quantitative-data/frequency-distribution-quantitative-data)

```{r}
head(faithful)
head(allDrugsNoNA)
durationF <- faithful$eruptions
range(durationF)
durationD <-  allDrugsNoNA$Number.of.Prescriptions
range(durationD)
breaksF <- seq(1.5, 5.5, by = 0.5)
breaksD <- seq(0, 40000, by = 500)
breaksD
duration.cutF <- cut(durationF, breaksF, right = F)
duration.cutD <- cut(durationD, breaksD, right = F)
options(scipen=999)
duration.freqF <- table(duration.cutF)
cbind(duration.freqF)

duration.freqD <- table(duration.cutD)
frequencyTableBind <- cbind(duration.freqD)
frequencyTableBind <- as.data.frame(frequencyTableBind)
dim(frequencyTableBind)
frequencyTableBind
frequencyTableBind["Bins"] <- seq(0, 39500, by = 500)
frequencyTableBind
```



