
---
title: "Medicaid Outpatient Narcotics"
output: html_notebook
---

This is an [Dataset](https://data.medicaid.gov/State-Drug-Utilization/State-Drug-Utilization-Data-2016/3v6v-qk5s) Notebook. When you execute code within the notebook, the results appear beneath the code. 

This is a [description of the dataset](file:///C:/Users/Chris%20Iyer/Downloads/Drug_Utilization_Data_Descriptions.pdf)

```{r}
library(knitr)
library(dplyr)
library(rvest)
library(ggplot2)
library(gridExtra)
library(descr)
library(treemap)
```

```{r}
medicaid <- read.csv("medicaid.csv", header = T, stringsAsFactors = F)

```

###Read in Data

```{r}

#treatment drug

CODEINE <- medicaid %>% filter(Product.Name == "TYLENOL W/")
CODEINE
    CODEINE1 <- medicaid %>% filter(Product.Name == "OXYCONTIN")
CODEINE1

```

###Medical Treatment

```{r}
SUBOXONE <- medicaid %>% filter(Product.Name == "SUBOXONE")
METHADONE <- medicaid %>% filter(Product.Name == "METHADONE")
NALTREXONE <- medicaid %>% filter(Product.Name =="NALTREXONE")
BUPRENORPHINE <-  medicaid %>% filter(Product.Name == "BUPRENORPHINE")
VIVITROL <- medicaid %>% filter(Product.Name == "VIVITROL")
ZUBSOLV <- medicaid %>% filter(Product.Name == "ZUBSOLV")
BUNAVAIL <- medicaid %>% filter(Product.Name == "BUNAVAIL")
TreatmentDrugs <- rbind.data.frame(BUNAVAIL, BUPRENORPHINE, METHADONE, NALTREXONE, SUBOXONE, VIVITROL, ZUBSOLV)
unique(TreatmentDrugs$Product.Name)
dim(TreatmentDrugs)
TREATMENT <- rbind.data.frame(SUBOXONE, METHADONE, NALTREXONE, BUPRENORPHINE, VIVITROL, ZUBSOLV, BUNAVAIL)
unique(TREATMENT$Product.Name)
dim(TREATMENT)
```


###Convert medicaid data to a dataframe

```{r}
Medicaid <- as.data.frame(medicaid)
```
###Retrieve narcotics list

```{r}
Narcotics <- read_html("https://www.drugs.com/drug-class/narcotic-analgesics.html")
Drug_Names <- Narcotics %>% html_nodes("br+ .small a") %>% html_text()


MORENARCS <- read_html("http://drugabuse.com/library/drugs-a-z/")

NARCNAMES <- MORENARCS %>% html_nodes("dt") %>% html_text() %>% gsub('[\r\n\t]', '', .)
NARCNAMES <- toupper(NARCNAMES)
ccc <- Medicaid %>% filter(Product.Name %in% NARCNAMES)

ccc <- ccc %>% filter(Product.Name != "CONCERTA" & Product.Name != "RITALIN" & Product.Name != "VIVITROL", Product.Name!= "SUBOXONE" & Product.Name!="METHADONE" & Product.Name!= "NALTREXONE" & Product.Name != "BUPRENORPHINE" & Product.Name != "ZUBSOLV" & Product.Name != "BUNAVIL" & Product.Name != "ADDERALL")

unique(ccc$Product.Name)
dim(ccc)
unique(TREATMENT$Product.Name)
```

Population Data

```{r}
Pops <- read_html("http://worldpopulationreview.com/states/")
StateName <- Pops %>% html_nodes("td:nth-child(2)") %>% html_text()
StatePopul <- Pops %>% html_nodes("td~ td+ td") %>% html_text()
StatePopul <- as.numeric(gsub(",", "", StatePopul))
Population <- cbind.data.frame(StateName, StatePopul)
Population
Population["State"] <- state.abb[match(Population$StateName,state.name)]
Population
```



###Filter some of the narcotics from the medicaid dataframe

```{r}
Drug_Names <- unique(toupper(Drug_Names))
Drug_Names
Drugs <- c("MORPHINE",
"OXYMORPHONE", 
"HYDROMORPHONE", 
"BUPRENORPHINE", 
"TRAMADOL", 
"HYDROCODONE", 
"BUTORPHANOL",  
"PROPOXYPHENE", 
"LEVORPHANOL",
"TAPENTADOL",
"PENTAZOCINE",  
"REMIFENTANIL", "OXYCONTIN")

medicaid1 <- Medicaid %>% filter(Product.Name %in% Drug_Names)
dim(medicaid1)
unique(medicaid1$Product.Name)
MORPHINE <- Medicaid %>% filter(grepl("MORPHINE", Product.Name))
OXYMORPHONE <- Medicaid %>% filter(grepl("OXYMORPHONE", Product.Name))
HYDROMORPHONE <- Medicaid %>% filter(grepl("HYDROMORPHONE", Product.Name))
BUPRENORPHINE <- Medicaid %>% filter(grepl("BUPRENORPHINE", Product.Name))
TRAMADOL <- Medicaid %>% filter(grepl("TRAMADOL", Product.Name))
HYDROCODONE <- Medicaid %>% filter(grepl("HYDROCODONE", Product.Name))
BUTORPHANOL  <- Medicaid %>% filter(grepl("BUTORPHANOL", Product.Name))
PROPOXYPHENE <- Medicaid %>% filter(grepl("PROPOXYPHENE", Product.Name))
LEVORPHANOL  <- Medicaid %>% filter(grepl("LEVORPHANOL", Product.Name)) 
TAPENTADOL <- Medicaid %>% filter(grepl("TAPENTADOL", Product.Name))
PENTAZOCINE <- Medicaid %>% filter(grepl("PENTAZOCINE", Product.Name))
REMIFENTANIL <- Medicaid %>% filter(grepl("REMIFENTANIL", Product.Name))
FIORICET <- Medicaid %>% filter(grepl("FIORICET", Product.Name))
TUSSINS <- Medicaid %>% filter(grepl("TUSS", Product.Name))
TUSSINS
toupper("Tylenol")
BUTALBITAL <- Medicaid %>% filter(grepl("BUTALBITAL", Product.Name))
# trazid <- Medicaid %>% filter(grepl("TRAZOD", Product.Name))
# trazid
```

###Bind narcotics into a dataframe

```{r}
allDrugs <- rbind.data.frame(medicaid1, MORPHINE, OXYMORPHONE, HYDROMORPHONE, BUPRENORPHINE, TRAMADOL, HYDROCODONE, BUTORPHANOL, PROPOXYPHENE, LEVORPHANOL, TAPENTADOL, PENTAZOCINE, REMIFENTANIL, FIORICET, TUSSINS, BUTALBITAL, CODEINE, ccc )
unique(allDrugs$Product.Name)

allDrugs %>% group_by(Product.Name) %>% summarise(count = n())
```

All the Narcotics in All States

```{r}
head(allDrugs)
allDrugs <- allDrugs %>% arrange(desc(Number.of.Prescriptions))
head(allDrugs, 35)
unique(allDrugs$Product.Name)
```

Percent of Rx in US that are Narcs

```{r}

percentNarcs <- dim(allDrugs %>% filter(State != "XX"))/dim(medicaid %>% filter(State != "XX"))
```



Percent of Rx in US that are used for Treatment of Opiod Addiction

```{r}
PercentRx <- dim(TREATMENT %>% filter(State != "XX"))/dim(medicaid %>% filter(State != "XX"))
PercentRx
```


Join Population with Treatment

```{r}
TREATMENT <- TREATMENT %>% select(State, Number.of.Prescriptions,Product.Name, Total.Amount.Reimbursed)

TreatPop <- TREATMENT %>% na.omit()
TreatPop <- TreatPop %>% group_by(State) %>% summarise(countRx = sum(Number.of.Prescriptions)) 
TreatPop <- inner_join(TreatPop, Population)
TreatPop <- TreatPop %>% arrange(StateName)


TreatPop1 <- cbind.data.frame(TreatPop, xxx)

TreatPop1 <- TreatPop1 %>% select(State, countRx, StatePopul, Difference)
TreatPop1 <- TreatPop1 %>% mutate(zRx = scale(countRx), zPop = scale(StatePopul), zDiff = scale(Difference))
PercentinTx <- TreatPop %>%  mutate(percentReceivingRx = round(countRx/StatePopul, digits = 3))

treemap(PercentinTx,index = c("percentReceivingRx","State"), vSize = "percentReceivingRx", vColor = "percentReceivingRx", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Treatment Rx per person")
```

```{r}
head(TREATMENT)
head(TreatPop)
head(xxx)
head(TreatPop1)
```


Opiod to Treatment Ratio

```{r}
percentNarcs/PercentRx
```


All drugs for this analysis

```{r}
allDrugs %>% group_by(Product.Name) %>% arrange(Product.Name) %>% summarise(count = n())
```


```{r}
a <- allDrugs %>% group_by(State) %>% summarise(entriesNA = sum(is.na(Number.of.Prescriptions)))
a <- a %>% filter(State != "XX")
a
b <- allDrugs %>% filter(State != "XX") %>% group_by(State) %>% summarise(entriesCount = n())
b
NAInfo <- inner_join(a,b)
c <- NAInfo %>% mutate(percentNA = round(entriesNA/entriesCount, digits = 2))
c
```

###Using a loop to make a seperate dataset for each state

```{r}
dim(allDrugs %>% filter(State != "XX"))
dim(allDrugs %>% filter(State !="XX") %>% na.omit())
#NA = 0.52
1-(28317/58439)

UniqueData <- allDrugs %>% filter(State != "XX")
unique(UniqueData$State)


for(i in unique(UniqueData$State)){
  
  
  newDS <- UniqueData %>% filter(State == i)
  
  print(head(newDS))
  hist(newDS$Number.of.Prescriptions, main = paste(i, "'s Rx Frequency"))
  identify(newDS$Number.of.Prescriptions)
  
}
```



```{r}
head(UniqueData)
head(c)
UniqueData <- inner_join(UniqueData, c)
head(UniqueData)
```


###There is so much missing data for each state, I would like to find a way to as accurately as possible calculate the number of prescriptions filled. Using a loop to make tables to show frequency of number of prescriptions for each state.

```{r}
breaksD <- seq(0, 45000, by = 500)


for(i in unique(UniqueData$State)){
  newDS <- UniqueData %>% filter(State == i)
  durationD <-  newDS$Number.of.Prescriptions
  duration.cutD <- cut(durationD, breaksD, right = F)
  options(scipen=999)
  duration.freqD <- table(duration.cutD)
  frequencyTableBind <- cbind(duration.freqD)
  frequencyTableBind <- as.data.frame(frequencyTableBind)
  frequencyTableBind[i] <- seq(0, 44500, by = 500)
  frequencyTableBind <- frequencyTableBind %>% mutate(relFreq = (duration.freqD/sum(duration.freqD)))
  frequencyTableBind$relFreq <- round(frequencyTableBind$relFreq, digits = 3)
  print(frequencyTableBind)
  
}
```






```{r}
breaksD <- seq(0, 45000, by = 500)
bins <- seq(1,91, by = 1)
bins
x <- cbind.data.frame(breaksD, bins)
x
```


```{r}
aSubset <- allDrugs %>% select(State,Number.of.Prescriptions)
head(aSubset)
aSubset <- aSubset %>% filter(State != "XX") %>% na.omit()
head(aSubset)

unique(aSubset$State)


```



Total number of prescriptions by state, NAs eliminated

```{r}
NAEliminated <- allDrugs %>% 
  group_by(State) %>% 
  summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE), sumAmount = sum(Medicaid.Amount.Reimbursed,na.rm=TRUE))
sum(NAEliminated$sumNumber)
NAEliminated

```

###Percent of number of Rx per state

```{r}
NAEliminated <-  NAEliminated[1:51, ]
for(i in 1:length(NAEliminated$State)){
  NAEliminated$percentofWhole <- (NAEliminated$sumNumber/sum(NAEliminated$sumNumber))
  NAEliminated$percentofWhole <- round(NAEliminated$percentofWhole, digits = 3)
}
NAEliminated$percentofWhole
```


Percent of NA Values per State, XX excluded

```{r}
numberNAValueState <- allDrugs %>% filter(State!= "XX") %>% group_by(State) %>% summarise(SumNA = sum(is.na(Number.of.Prescriptions)))
head(numberNAValueState)
sum(numberNAValueState$SumNA)
```

###Percent of state entries that are NA and state's NA percentage relative to the USA

```{r}

sum(is.na(allDrugs))
allDrugsXX <- allDrugs %>% filter(State != "XX") %>% group_by(State) %>% summarise(percentOfStateEntriesThatAreNA = mean(is.na(Number.of.Prescriptions)), numberNA = sum(is.na(Number.of.Prescriptions)), PercentofAllNA = numberNA/sum(numberNAValueState$SumNA))

head(allDrugsXX)
allDrugsXX <- allDrugsXX %>% mutate_each(funs(round(.,3)), -State) 
allDrugsXX

```


State reimbursement for Drugs relative to the country, pie in ggplot

```{r}
statePortion <- NAEliminated %>% mutate(portion = sumAmount/sum(sumAmount))
statePortion

bp <- ggplot(statePortion, aes(x="", y=portion, fill=State, color = "red")) + geom_bar(width = 1, stat = "identity")
bp + coord_polar("y", start=0)
```

State reimbursement for Drugs relative to the country, pie in plotrix

```{r, fig.height=10, fig.width=10}
library(plotrix)
pie3D(statePortion$portion, labels = statePortion$State, main = "State Portion of Spending", explode=0.1, radius=.9, labelcex = 0.8,  start=0.7)
```


```{r, fig.height=10, fig.width=10}

head(allDrugsXX)
treemap(allDrugsXX,index = c("PercentofAllNA","State"), vSize = "PercentofAllNA", vColor = "PercentofAllNA", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Each State Percent of NA \nRelative to Country")

treemap(allDrugsXX,index = c("percentOfStateEntriesThatAreNA", "State"), vSize = "percentOfStateEntriesThatAreNA", vColor = "percentOfStateEntriesThatAreNA", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Percent of States' Entries that are NA")
```



All Drugs with State XX removed, NAs eliminated

```{r}
#Number of Rx and amount reimbursed per state
States <- allDrugs %>% 
  group_by(State) %>% 
  summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE), sumAmount = sum(Medicaid.Amount.Reimbursed,na.rm=TRUE), numberEntries = n())
States
```
NA counts. percentages, and sums of Rx: Join

```{r}

Join <- inner_join(States, allDrugsXX) 

```


###Join drug reimbursement with 2016 population: allDrugs

```{r}
allDrugs <- inner_join(allDrugs, Population)

dim(allDrugs)
```

###Join State Rx filled with Population: allDrugsnoXX

```{r}
allDrugsnoXX <- inner_join(States, Population)
dim(allDrugsnoXX)
```


###Overall Average of Rx filled per capita in USA

```{r}
averageRxPP <- allDrugsnoXX %>% summarise(AverageRxPPinUSA = sum(sumNumber)/sum(StatePopul))
averageRxPP
```

```{r, fig.height=4, fig.width=12}

averageRxPP

StateAveRxPP <- allDrugsnoXX %>% mutate(RxPP = sumNumber/StatePopul) %>% mutate_each(funs(round(.,4)), -c(State, StateName))

StateAveRxPP <- StateAveRxPP %>% mutate(Status = ifelse(RxPP <= averageRxPP$AverageRxPPinUSA, "Below Average", "Above Average"))

StateAveRxPP
PlotA <- ggplot(StateAveRxPP, aes(x = State, y = RxPP * 100000, fill = Status)) + geom_bar(stat = "identity") + 
    geom_hline(yintercept = 0.1246144 * 100000	) + 
scale_fill_manual(values = c("purple", "blue")) + ggtitle("State average Rx per 100,000 as compared to US average") + 
     theme(plot.title = element_text(lineheight=.8, face="bold")) +
  scale_y_continuous("Opiod Rx per 100,000") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r}
RxPerCap <- treemap(StateAveRxPP,index = c("RxPP", "State"), vSize = "RxPP", vColor = "RxPP", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Number of Rx Per Capita")
```

###Missing Data: Pcecent NA Values per State

```{r}
treemap(allDrugsXX,index = c("percentOfStateEntriesThatAreNA", "State"), vSize = "percentOfStateEntriesThatAreNA", vColor = "percentOfStateEntriesThatAreNA", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Percent of missing entries")

```

Number of prescriptions per capita, slightly different code

```{r}
PerPerson <- allDrugsnoXX %>% mutate(RxPP = sumNumber/StatePopul, RxSpendingPP = sumAmount/StatePopul) %>% arrange(desc(RxPP))
PerPerson
PerPerson <- PerPerson %>% mutate_each(funs(round(.,3)), -c(State, StateName))
PrescriptionsPerPerson <- treemap(PerPerson,index = c("RxPP", "State"), vSize = "RxPP", vColor = "RxPP", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Rx Per Person")

```


[Opiod Deaths in 2015](https://www.cdc.gov/drugoverdose/data/statedeaths.html)

```{r}
OpiodDeath <- read.csv("OpiodDeath.csv", header = T)
OpiodDeath <- OpiodDeath[c(1:51), c(1,4)]
OpiodDeath$Number <- as.numeric(gsub(",", "", OpiodDeath$Number))
OpiodDeath 
```


Opiod Death per capita

```{r}
DeathPerCap <- inner_join(PerPerson, OpiodDeath)
DeathPerCap <-  DeathPerCap %>% mutate(ODPP = Number/StatePopul)
DeathPerCap <- DeathPerCap%>% mutate_each(funs(round(.,3)), -c(State,StateName, ODPP )) %>% mutate_each(funs(round(.,6)), c(ODPP))

DeathPerCap
Death <- treemap(DeathPerCap,index = c("ODPP", "State"), vSize = "ODPP", vColor = "ODPP", 
         align.labels = list(c("left", "top"), c("right", "bottom")), title = "Number Deaths Per Capita")

OverallDeathperCapUSA <- sum(DeathPerCap$Number)/sum(DeathPerCap$StatePopul)
OverallDeathperCapUSA

DeathPerCap <- DeathPerCap %>% mutate(DeathPosition = ifelse(ODPP > OverallDeathperCapUSA, "Above Average", "Below Average"))
DeathPerCap

PlotB <- ggplot(DeathPerCap, aes(x = State, y = ODPP * 100000, fill = DeathPosition)) + geom_bar(stat = "identity") + 
    geom_hline(yintercept = 0.0001604907 * 100000) + 
scale_fill_manual(values = c("purple", "blue")) + ggtitle("State average opiod related per 100,000 as compared to US average") + 
     theme(plot.title = element_text(lineheight=.8, face="bold")) +
  scale_y_continuous("Overdose deaths per 100,000") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r, fig.height=3, fig.width=10}
grid.arrange(PlotA, PlotB, ncol = 1)
```

Z Score: where you subtract the mean and divide by the standard deviation of your variable

```{r}
DeathPerCap

meanOD <- mean(DeathPerCap$ODPP)
meanOD
sdOD <- sd(DeathPerCap$ODPP)

sdOD

```

```{r}
meanRx <- mean(DeathPerCap$RxPP)
meanRx
sdRx <- sd(DeathPerCap$RxPP)
sdRx
```


```{r}
names(DeathPerCap)
DeathandRxperCapperState <- DeathPerCap 
names(DeathandRxperCapperState)
DeathandRxperCapperState <- DeathandRxperCapperState %>% mutate(zRx = ((RxPP - mean(RxPP))/ sd(RxPP)), zOD = ((ODPP - mean(ODPP))/sd(ODPP)))
DeathandRxperCapperState
```
```{r, fig.height=4, fig.width=10}
names(DeathandRxperCapperState)
library(reshape2)
newset <- DeathandRxperCapperState %>% select(State, zRx, zOD)
newset
dat.m <- melt(newset)
range(dat.m$value)
dat.m

newset

ggplot(dat.m, aes(State, value, fill=variable)) + 
  geom_bar(position="dodge",stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_manual(values = c("red", "blue")) 

head(TreatPop1, 1)
Melted <- TreatPop1 %>% select(State, zRx, zPop, zDiff)
anotherMelted <- melt(Melted)
anotherMelted


ggplot(anotherMelted, aes(State, value, fill=variable)) + 
  geom_bar(position="dodge",stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_manual(values = c("red", "blue", "green")) 

```

```{r}
names(DeathandRxperCapperState)
forCorr <- DeathandRxperCapperState %>% select(State, zRx, zOD)
```

```{r}
library(Hmisc)
head(forCorr, 1)
m1<-lm(zOD ~ zRx,data=forCorr)
summary(m1)
forCorr <- forCorr %>% arrange(State)
head(Melted, 1)

cc <- cbind.data.frame(Melted, forCorr)
cc <- cc[ , c(1:2, 4,7)]
cc
head(TreatPop1)
m <- inner_join(OpiodDeath, TreatPop1) 
m <- m[,c(1:5)]
m

```

```{r}
plot1 <- ggplot(forCorr, aes(x=zRx, y=zOD)) +
    geom_point(shape=1)      # Use hollow circles
```

```{r}
plot1 +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm)   # Add linear regression line 
                             #  (by default includes 95% confidence region)
```

```{r}
plot1 + geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm,   # Add linear regression line
                se=FALSE)    # Don't add shaded confidence region
```

```{r}
plot1 + geom_point(shape=1) + geom_smooth()
```

```{r}
plot1 + geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) + geom_point()
```


Quarterly number of prescriptions

```{r}
head(allDrugs)
Quarterly <- allDrugs %>% group_by(State, Quarter) %>% summarise(sumNumber = sum(Number.of.Prescriptions,na.rm=TRUE))
```

###Quarterly Count of NA Values per state
We don't know the number of prescriptions for each NA value

```{r}
Subs1<-subset(allDrugs, (is.na(allDrugs[,11]))) 
#Total number of NA
dim(Subs1)
head(Subs1)
```

###Quartrely NA for states

```{r}
Subs2 <- Subs1%>% group_by(State, Quarter) %>% summarise(NA_count = n())
dim(Subs2)
head(Subs2, 28)
```

#Number of NA for each state in 2016

```{r}
Subs3 <- Subs1 %>% group_by(State) %>% summarise(NA_count = n())
dim(Subs3)
kable(Subs3)
```


Percent of NA entries in the allDrugs dataset. 46%

```{r}
NoNA <- allDrugs %>% na.omit() 
head(NoNA)

dim(allDrugs)
dim(NoNA)
1-(101683/188511)
```




###ALASKA

```{r}
AK <- allDrugs %>% filter(State == "AK") 
dim(AK)
#There are 826 entries in AK
sum(is.na(AK$Number.of.Prescriptions))
summary(AK)
  
head(AK)
#Percentage of values that are NA
1374/2524
hist(AK$Number.of.Prescriptions)
```

###Percent NA in Maine

```{r}
ME <- allDrugs %>% filter(State == "ME") 
dim(ME)
sum(is.na(ME$Number.of.Prescriptions))
summary(ME)
head(ME)
hist(ME$Number.of.Prescriptions)
```

###CA

```{r}
CA <- allDrugs %>% filter(State == "CA") 
dim(CA)
sum(is.na(CA$Number.of.Prescriptions))
summary(CA)
head(CA)

hist(CA$Number.of.Prescriptions)
```


Visualization of missing values

```{r, fig.height=8, fig.width=10}
library(Amelia)

missmap(allDrugs, main="Missing Data Map")
```


```{r}
library(reshape2)
library(ggplot2)
library(plyr)
allDrugsNoNA <- allDrugs %>% na.omit()
head(allDrugsNoNA)
dim(allDrugsNoNA)
summary(allDrugs)
allDrugsNoNA$z <- ave(allDrugsNoNA$Number.of.Prescriptions, allDrugsNoNA$State, FUN=scale)
head(allDrugsNoNA)

new_DF <- allDrugs[rowSums(is.na(allDrugs)) > 0,]
new_DF[new_DF=='NA'] <- NA
head(new_DF)


```

```{r, eval=FALSE, fig.height=75, fig.width=10, include=FALSE}

ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions, fill = factor(State))) + geom_histogram(binwidth = 150) + stat_bin(binwidth = 100, geom="text", aes(label=..count..), angle = 90) + facet_wrap(~ State, ncol = 1) +  theme_bw() + scale_x_continuous(limits = c(0, 5000)) 

```

[labels on ggplot](https://stackoverflow.com/questions/30057765/histogram-ggplot-show-count-label-for-each-bin-for-each-category)

```{r, fig.height=45, fig.width=10}

breaksD
#head(allDrugsNoNA)
ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions, group= factor(State), fill = "white")) + geom_histogram() + stat_bin(binwidth = 500, geom="text", aes(label=..count..), angle = 90) + facet_wrap(~ State, ncol = 2) + theme_bw() + scale_x_continuous(limits = c(-100, 1100), breaks = breaksD) +
  theme(axis.text.x=element_text(angle=90,hjust=1)) 
```


```{r, fig.height=45, fig.width=10}
p7 <- ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions)) +
        geom_histogram() + facet_wrap(~State, ncol = 2) + scale_x_continuous(limits = c(-100, 10000), breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000, 11000)) + scale_y_continuous(limits = c(-100,275)) + stat_bin(binwidth = 400, geom="text", aes(label=..count..), angle = 90)
p7
```


```{r}
library(dplyr)
sum(is.na(allDrugs$Number.of.Prescriptions))



```


```{r, fig.height=56, fig.width=10}
ggplot(allDrugsNoNA,aes(x=Number.of.Prescriptions)) + 
    stat_bin(binwidth=500, geom="text", aes(label=..count..), vjust=-1.5, angle = 90) + ylim(c(-100, 550)) +  facet_wrap(~State, ncol = 1) + scale_x_continuous(limits = c(-2000, 6000), breaks = c(-2000,-1000, 0, 500,1000, 1500,2000, 2500,3000, 3500,4000, 4500,5000) ) +
  theme(axis.text.x=element_text(angle=90,hjust=1)) 
```



```{r, fig.height=50, fig.width=10}
dim(allDrugsNoNA)
head(allDrugsNoNA)
ggplot(allDrugsNoNA, aes(x = Number.of.Prescriptions )) + 
  geom_histogram(aes(fill = ..count..), breaks = seq(0,1000, by = 100),binwidth=100, colour="blue", fill = "blue", alpha = 0.2, lwd=0.2) +
  stat_bin(aes(label = ..count..),geom="text", colour="black", vjust=-1, size=4.5, breaks = c(0, 100, 200,300,400,500, 600,700,800,900,1000)) +
  scale_x_continuous(limits = c(0, 1000),labels = c(0, 100, 200,300,400,500, 600,700,800,900,1000), breaks = c(0, 100, 200,300,400,500, 600,700,800,900,1000)) + scale_y_continuous(limits = c(0,500), labels = c(0, 100, 200, 300, 400, 500))+ facet_wrap(~State, ncol = 2) + theme(axis.text.x=element_text(angle=90,hjust=0)) 
```




[How to do a frequency table in base R](http://www.r-tutor.com/elementary-statistics/quantitative-data/frequency-distribution-quantitative-data)

```{r}
head(faithful)
durationF <- faithful$eruptions
range(durationF)
breaksF <- seq(1.5, 5.5, by = 0.5)
duration.cutF <- cut(durationF, breaksF, right = F)
duration.freqF <- table(duration.cutF)
cbind(duration.freqF)
```




```{r}


allDrugsNoNA <- allDrugsNoNA %>% arrange(desc(State))
allDrugs <- allDrugs %>% arrange(desc(State))
head(allDrugsNoNA)
head(allDrugs)
dim(allDrugs)
dim(allDrugsNoNA)

```




```{r}
dim(allDrugsNoNA)
allDrugsNoNA %>% arrange(desc(State))
durationD <-  allDrugsNoNA$Number.of.Prescriptions
range(durationD)
breaksD <- seq(0, 45000, by = 500)
breaksD
duration.cutD <- cut(durationD, breaksD, right = F)
options(scipen=999)
duration.freqD <- table(duration.cutD)
frequencyTableBind <- cbind(duration.freqD)
frequencyTableBind <- as.data.frame(frequencyTableBind)
dim(frequencyTableBind)
frequencyTableBind
frequencyTableBind["Bins"] <- seq(0, 44500, by = 500)



frequencyTableBind %>% mutate(averageBin = cummean(Bins), averageMid = cummean(Bins)*duration.freqD)
```





###

```{r}
Binned <- frequencyTableBind %>% mutate(midpt = Bins + 250,  percentofWhole = duration.freqD/sum(duration.freqD))
Binned
sum(Binned$duration.freqD)
```

```{r}
head(allDrugsNoNA) %>% filter(State != "XX")
allDrugsNoNA %>% filter(State != "XX") %>% summarise(sum(Number.of.Prescriptions))

```

```{r}


head(allDrugsNoNA)
sum(allDrugsNoNA$Number.of.Prescriptions)

head(Binned)

```


```{r}
allDrugsXX
head(allDrugsNoNA)

tryThis <- inner_join(allDrugsXX, allDrugsNoNA)
tail(tryThis)
dim(tryThis)
frequencyTableBind

allDrugsNoNA %>% group_by(State) %>% summarise(meanNRx = mean(Number.of.Prescriptions))
```




```{r}

library(plyr)
count(tryThis, 'State')
count(tryThis, 'Number.of.Prescriptions')
tryThis %>% group_by(State,Number.of.Prescriptions) %>% select(State, Number.of.Prescriptions) %>% tally()
```

CA freq dist

```{r}
durationD <-  CA$Number.of.Prescriptions
range(durationD)
breaksD <- seq(0, 45000, by = 500)
breaksD
duration.cutD <- cut(durationD, breaksD, right = F)
options(scipen=999)
duration.freqD <- table(duration.cutD)
frequencyTableBind <- cbind(duration.freqD)
frequencyTableBind <- as.data.frame(frequencyTableBind)
dim(frequencyTableBind)
frequencyTableBind
frequencyTableBind["Bins"] <- seq(0, 44500, by = 500)

frequencyTableBind %>% mutate(averageBin = cummean(Bins), midpoint = averageBin+250, Numbermissing = duration.freqD*midpoint*0.439) %>% select(duration.freqD, Bins, midpoint, Numbermissing)
#total = 1629
sum(frequencyTableBind$duration.freqD)


frequencyTable <- frequencyTableBind %>% mutate(midpt = cummean(Bins) + 250, relFreq = duration.freqD/1629, missing = .43*duration.freqD, sumRxMissing = missing*midpt)
frequencyTable
xYZCA <- CA %>% filter(Number.of.Prescriptions > 500 & Number.of.Prescriptions <= 1000)
xYZCA
mean(xYZCA$Number.of.Prescriptions)
```



```{r}
iris_split <- split(iris, iris$Species)
new_names <- c("one", "two", "three")
for (i in 1:length(iris_split)) {
  assign(new_names[i], iris_split[[i]])
}
one
```


```{r}
head(allDrugsNoNA)
unique(allDrugsNoNA$State)
states_split <- split(allDrugsNoNA, allDrugsNoNA$State)
new_names <- sort(state.abb)


for (i in 1:length(states_split)){
  
  assign(new_names[i], states_split[[i]] %>% select(State, Number.of.Prescriptions))
  

}
WY

#Population["State"] <- state.abb[match(Population$StateName,state.name)]
```

```{r}
br = seq(0,45000,by=5000)
ranges = paste(head(br,-1), br[-1], sep=" - ")
my_bin <- function(x){
  


freq   = hist(x, breaks=br, include.lowest=TRUE, plot=TRUE)
#data.frame(freq)
summary(x)

#as.data.frame(range = ranges, frequency = freq$counts)
}




my_bin(CA)
```

```{r}
x <- cbind( Freq=table(WY$Number.of.Prescriptions), Cumul=cumsum(table(WY$Number.of.Prescriptions)), relative=prop.table(table(WY$Number.of.Prescriptions)))
x <- data.frame(x)
x


```



